\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-02 coqimpl.tex
Going through before metatheory for small-fixes.

\subsection{Language Design}\YZ{I think in §3.1, you should use a figure to show how the actual code the programmer will write to solve the programming challenge outlined in Figure 1.}
Inspired by our above challenges and Family Polymorphism, we propose
the following extension to Coq's Vernacular Commands.
% basically write down the specification of surface Coq commandline
%   and then explain  how the language design makes the programmer’s life easier
\begin{align*}
  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
  \quad |\quad "EndFamily" \\
  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
  & | \quad "FTheorem"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle \\
  & |\quad "Extend"~"Family"~\langle ident\rangle \\ 
  & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
  & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad | \quad ...
\end{align*}

@@

 Extend FInductive term : Set := 
  | tt : term | ff : term 
  | tif : term -> term -> term -> term
 Extend Family subst_internal.
  Final Field tt := ..
  Final Field ff := ..
  Final Field tif := ..
 EndFamily.
 Inherit subst.
(* Plugin will redo exhaustiveness check, 
  in case subst_internal is not 
  extended correctly *)
 Extend FInductive has_type (* .. and new ty too *)
  : context -> term -> type := ...
 Extend FTheorem subst_lemma.
  ... (* Need to prove extra cases *)
 Qed.

@@

These are some core extensions we proposed to support \cref{fig:STLC-example}. We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family. "FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
We can define nested families.

To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness. The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family. Then
"FRecursor" will be used to construct a recursor using the handler family.
For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
"handlers" will generate a new field "recur" that is doing induction
(using elimination principle of "T") using motive "f", and the handlers
are located in the family "handlers". This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family. "FTheorem" is used for theorem proving

We can inherit a top-level family "parent" using \mintinline{Coq}{Family
... extends parent} as a new top-level family. As expected, "Override"
will override a field from the parent. "Inherit" can be used to inherit
nested families, inductive types, and fields. The fields inside a family
are order-sensitive because we have dependent types here. So "Inherit
Until" is used if the user wants to inherit a range of fields. If the
user wants to extend an inductive type or a nested family from the
parent, then "Extend FInductive" or "Extend Family" is used. Thus nested
family can only be inherited or extended together with its outer family
(and thus top level family), thus our support of family polymorphism is
incomplete.

@@

After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.

@@

\textbf{Family compiled into module.} We support family by compiling
family components into Coq modules, family ``types'' into module types, and the context of a given term
into parameters of the module. We need to faithfully
reflect the \textit{late binding} nature of families: for each defining
field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
choose module and module parameters to achieve this wrapping---each
family type in the context of the judgement will be one module parameter
with the compiled module type, and thus each field of the family will
compile into a functor parametrized by its context. Doing so we can also
get rapid feedback from type checking when the users are defining each
field interactively with Coq.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}[!htb]
  \begin{minipage}[t]{0.30\linewidth}
\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
Family A.
Field a : nat := 1.
Field b : a = a := eq_refl.
Final Field a' : nat := 1. 
EndFamily.
\end{lstlisting}
  \end{minipage}%\YZ{Can the self's be omitted?}\EDJreply{Yes. But in our current plugin implementation it is not omitted. I think altering example is fine. If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}\EDJreply{Omitting self_ is easy and not much engineering effort is required. But omitting self_ will lead to `A.a' instead of `a' actually. So if we really want `a' as you last time mentioned, then a resolution is required and more engineering effort is required. But I don't see any technical difficulties. I have removed all the self, please check.}
  \begin{minipage}[t]{0.35\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_5 (self_A: EmptySig).
Parameter (a : nat). End a_5.
Module Type a_6 := a_5 EmptyMod.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = self_A.a 
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: when the user
interactively inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_5, a_6"
three components.
{The "EmptySig" is just an empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_4" is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking as in \ref{chg:software-engineering}; overriding is also directly supported---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. This compiled module type serves as an
abstraction that enables overriding.

@@

% explaining inheritance
\textbf{Implementing Inheritance.}
To resolve \ref{chg:consistency}, we make sure when inheritance happens, not only the content will be inherited, but the context of the corresponding field is also ``inherited''. Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty and has no "g" in "self".

@@

There are three kinds of inheritance---extend, override, and inherit. We
use a list to encode inheritance data inside the
implementation of the plugin. The inheritance data will include the
information of (1) the ``type'' of the parent family, (2) a list of
operation indicating how each field of the parent will be dealt with
(either inherited or overriden), (3) the operations indicating newly
extended fields, and (4) the ``type'' of the children family.
The surface syntax for the three kinds of inheritance
are demonstrated in the exemplar \cref*{fig:plugin-example3}. The
implementation of overriding, inheritance, and extension is achieved by
simply manipulating the internal representation of a family
correspondingly---we can swap the module in the list for overriding,
retain the module in the list for inheritance, and add new module into
the list for extension. The compilation from a family (internal
representation) to a module still acts the same.

Though inheritance is solely a first-order data rather
than a higher-order functor, it is still possible to
\text{mix-in} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance. However, a good definition of mixin is still under investigation. We will show one practical example of \textit{mixin} later.\YZ{How about making mixin composition one of your Req's? Also does any example in the paper illustrate how to do this?}\EDJreply{mixin is not really formalized. There is no example illustrate it. The only example is the STLC example at the very end and I just kind of ``mention'' it. I think here I just want to emphasize, inheritance as first order data is still possible to support composition.}

@@

% For example, we may
% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
% and intend to prove its commutativity in a following field $"comm"$.
% But to prove so, we need to know more than just a type interface of
% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
% future, once "add" is overridden into a subtraction function then
% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
% A more concrete example is in \cref{fig:plugin-example1}, consider the
% case we have "Field b : a = 1 := .." instead. Then our
% compilation and Coq's type checking will both fail: in the context of
% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
% more (and this is necessary so that "a" can be overridden in the
% future).
Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the
example. We can see that the first attempt of proving "a = 1" fails.
\textit{This should fail}---because we want to retain the possibility
to override "a" in the future, but the mere existence of the field "c"
will stop that, since "c" witnesses "a" fix to "1". In our implementation, \textit{this will fail}---the proof of "c" will require a concrete definition of "a",
but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in the context where we only know \mintinline{Coq}{a : nat} thus failed.

@@

\textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.} 
Non-overridable fields are still not expressive enough to achieve the full
extent of dependent type programming. There are cases where overriding and
(definition-relevant) reasoning are both needed. For example, we may want a field to be an overridable monotonic function. We cannot split into two fields
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
because to prove
the latter we need to have concrete definition of "f" in the family type, which
makes "f" non-overridable at all. One option is to use sigma type
\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
a bundle of stuff with multiple refinements.

To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type. In a sealed family, every field will expose its definition inside the family but a family as a whole will be sealed into a relaxed type. Thus we can override one sealed family as a whole and thus change the internal implementation.

@@

And the compiled module type of
"MonotonicF" will be fixed throughout the extension in different
families. Overriding field will always have these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }

In our example, the seal family "I.MonotonicF" is actually simulating
interfaces as well.  We have
\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
and we have assigned it to both "f" and "p", and later instantiate in
"I2" with concrete implementation by overriding this family
"I.MonotonicF".
% Basically, {a : int -> int; property : a is monotonic}, in this case
%   we may want to override "a" using different computation, but we still wnat
%   this refinement/property
The implementation of sealed family is not so different from that of the normal family, but we need to be careful about the compiled module type.

@@

We start with the programming interface. We use "FInductive" to define a extensible inductive type and in the children family we use "Extend Finductive ..." to extend the inductive type with new constructors. In \cref{fig:plugin-example2}, we make the example of boolean and three-valued boolean.

Recall, to construct a recursor, we need to create for each
constructor a handler (function) and aggregate them into a family, and thus that family (e.g. "neg_handler" in \cref{fig:plugin-example2}) can be considered as a large matching-case statement. We use "FRecursor" to declare a recursor (e.g. "neg") by specifying the motive, the aiming inductive type and the handler family. Once recursor is created it will look like any other function fields (e.g. "neg"). Here in \cref{fig:plugin-example2} we apply "neg" to "tt" as an exemplar field. The inheritance of recursor is mainly delegated by the inheritance of the handler family---if the user extend with new constructor (e.g. "uu"), then the inherited family (e.g. "neg_handler") will need to extend correspondingly (e.g. "neg_handler.uu"). If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".

@@

Let's look at the exemplar \cref{fig:plugin-example2}, the example of boolean and negation. When we define the extensible inductive type "b", all the following fields (e.g. "neg_handler.ff") can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface''---the compiled module type "b_3". We cannot export\YZ{export or put?}\EDJreply{Ok. I just figure out how my wording is confusing. put is absolutely correct but by put I want to mean ``export''. I think I use ``export'' in several places...}
the eliminator of the inductive type "b" inside "b_3" because doing so all of the following fields (e.g. "neg_handler.ff" and the corresponding "ff_24") will not be able to be inherited to the context where "b"
is extended with a third constructor, \textbf{since the existence of recursor
alone will assert that there are only two constructors} in "b", not
\textbf{extensible} at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}\EDJreply{You are right. It is not only about "ff_24".I rephrased it. Please check.}

However, we later still need the concrete eliminator "b" for doing type-checking (as in \ref{chg:extensible-inductive-type}), thus the information doesn't totally disappear but store elsewhere.  Here we show one exemplar extraction of the recursor to \mintinline{Coq}{Set} as module "b_rec_12". This concrete recursor will be used by our "FRecursor". For example, in \cref{fig:plugin-example2}, "FRecursor" command will feed Coq a module "v_33_34" for doing type-checking with specified handlers (e.g. "neg_handler") during compilation---what it does is basically
assert the existence of the recursor and apply to see if handlers can
pass the type checking by Coq. Note that, at this point of the example, "FRecursor" will only do type check. The compilation of "FRecursor" into a component of the compiled module only happens when closing the whole family, by inserting the following into the compiled module: 
\begin{minted}{Coq}
  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff.
\end{minted}
where "b_rec" and "b" are the corresponding recursor and the (vanilla) inductive type in that compiled module. Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields.

Importantly, the recursor field (e.g. "neg") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) module, a recursor field is not compiled like that. For all the children family, we have to reconstruct and override with a ``new'' recursor using the (maybe inherited) handler family. This is affordable since each recursor handler doesn't need
to be rechecked---look at "B2.neg_handler", there is only one new
handler for "uu" and the handlers (mundane fields) from the parent
"B.neg_handler.tt,B.neg_handler.ff" have already been type-checked and
inherited (since they only rely on ``the abstracted
interface'' module "b_3"). 
Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg". For example, for the new "neg", when compiling family "B2" into a module, at the appropriate place our plugin will insert
\begin{minted}{Coq}
  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff neg_handler.uu.
\end{minted}
where "b_rec" and "b" are the corresponding recursor and the (vanilla but extended and modified) inductive type in that compiled module. Note that, even though "b_rec" and "b" has the same name as before, they are in the newly-compiled module so they refer to totally different things.

Of course, in the implementation, our plugin will ``inherit'' the
recursor "neg" in the surface \YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}\EDJreply{Added above. Please Check.}
but what the plugin actually does is doing a second exhaustiveness
checking in children family like mentioned above, with the ``same'' motive and ``same'' handler family.\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea. Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}\EDJreply{Please check.}

@@

% explain ftheorem wrapping this complicated recursor
This design of decoupling of the case handlers and exhaustiveness checking handles the semantic well, however, it brings a lot of boilerplate code---we need to create an extra (handler) family, manually specifying the type of each case handlers, and then using "FRecursor" to ``tie the knot'' and construct a recursor field from the family.  
We thus provide a command "FTheorem" to remove these boilerplate in programming interface when using tactic programming instead. 
This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor".\YZ{example?}\EDJreply{FTheorem is really just a short hand for "FRecursor" + handler family. Nothing more conceptual here. If the reader is interested in the concrete syntax, I think I need to make the two examples avaliable in the appendix.} We expect this "FTheorem" to be used in theorem proving but not general programming.

>>>>

\subsection{Language Design}\YZ{I think in §3.1, you should use a figure to show how the actual code the programmer will write to solve the programming challenge outlined in Figure 1.}\EDJreply{Done.}
Inspired by the above challenges and Family Polymorphism, we propose
the following extension to Coq's Vernacular Commands, and the corresponding STLC example.
% basically write down the specification of surface Coq commandline
%   and then explain  how the language design makes the programmer’s life easier
\begin{align*}
  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
  \quad |\quad "EndFamily" \\
  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
  & | \quad "FTheorem"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle \\
  & |\quad "Extend"~"Family"~\langle ident\rangle \\ 
  & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
  & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad | \quad ...
\end{align*}

@@

 Extend FInductive term : Set := 
  | tt : term | ff : term 
  | tif : term -> term -> term -> term
 Extend Family subst_internal.
  Final Field tt := ..
  Final Field ff := ..
  Final Field tif := ..
 EndFamily. (* Automatic inheriting others 
              from STLC.subst_internal *)
 Inherit subst.
(* Plugin will redo exhaustiveness check, in case 
 subst_internal is not extended correctly *)
 Extend FInductive has_type (* .. and new ty too *)
  : context -> term -> type := ...
 Extend FTheorem subst_lemma.
  ... (* Need to prove extra cases *)
 Qed.

@@

These are the core extensions we proposed to support \cref{fig:STLC-example}. We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family. "FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
Nested families are also allowed.

To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness. The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family. Then
"FRecursor" will be used to construct a recursor by refering to the handler family.
\mintinline{Coq}{FRecursor recur about T motive f } "using"
"handlers" will generate a new field "recur" that is really doing computation
(using elimination principle of the inductive type "T") with motive "f". The handlers
are located in the family "handlers", similar to a large pattern-matching statement. This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family. For example, in \cref{fig:STLC-example2},  "FTheorem" has a similar functionality but is expected to be used for theorem proving---the handler and the family will be automatically constructed during proving.

We can inherit a top-level family "parent" using \mintinline{Coq}{Family
... extends parent} as a new top-level family. As expected, "Override"
will override a field from the parent. "Inherit" can be used to inherit
nested families, inductive types, and fields. The fields inside a family
are order-sensitive because we have dependent types here. So "Inherit
Until" is used if the user wants to inherit a range of fields.  "Extend Family" is used only inside a inheriting family. Thus nested
family can only be inherited or extended together with its belonging family
(i.e. ultimately top level family), thus our support of family polymorphism is
incomplete.

@@

After considering the pros and cons, we decide to implement a Coq Plugin where we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the orignal surface Vernacular command) on the fly, instead of modifying the code-base of Coq.

@@

\textbf{Family compiled into module.} The main idea of our plugin is to support family by compiling
family components into Coq modules, family ``types'' into module types, and the context of a given term
into parameters of the module. We need to faithfully
reflect the \textit{late binding} nature of families: for each defining
field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
choose module and module parameters to achieve this wrapping---each
family type in the context will be one module parameter
with the compiled module type, and thus each field of the family will
compile into a functor parametrized by its context. Doing so we can also
get rapid feedback from type checking when the users are defining each
field interactively with Coq.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}[!htb]
  \begin{minipage}[t]{0.30\linewidth}
\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
Family A.
Field a : nat := 1.
Field b : a = a := eq_refl.
Final Field a' : nat := 1. 
EndFamily.
\end{lstlisting}
  \end{minipage}%\YZ{Can the self's be omitted?}\EDJreply{Yes. But in our current plugin implementation it is not omitted. I think altering example is fine. If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}\EDJreply{Omitting self_ is easy and not much engineering effort is required. But omitting self_ will lead to `A.a' instead of `a' actually. So if we really want `a' as you last time mentioned, then a resolution is required and more engineering effort is required. But I don't see any technical difficulties. I have removed all the self, please check.}
  \begin{minipage}[t]{0.35\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_5 (self_A: EmptySig).
Parameter (a : nat). End a_5.
Module Type a_6 := a_5 EmptyMod.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = self_A.a 
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: immediately after the user inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_5, a_6"
three components.
{The "EmptySig" is just an empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_4" is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking as in \ref{chg:software-engineering}; overriding is also directly supported---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. This compiled module type serves as an
abstraction that enables overriding.

@@

% explaining inheritance
\textbf{Implementing Inheritance.}
To resolve \ref{chg:consistency}, we make sure when inheritance happens, not only the content will be inherited, but the context of the corresponding field is also ``inherited''. Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty thus "self.g" doesn't type-check.

@@

There are three kinds of inheritance---extend, override, and inherit. Simiarly, we
use a list to encode inheritance data in the
implementation. The inheritance data will include the
information of (1) the ``type'' of the parent family, (2) a list of
operation indicating how each field of the parent will be dealt with
(either inherited or overriden), (3) the operations indicating newly
extended fields, and (4) the ``type'' of the children family.
The surface syntax for the three kinds of inheritance
are demonstrated in the exemplar \cref*{fig:plugin-example3}. The
implementation of overriding, inheritance, and extension is achieved by
simply manipulating the internal representation of a family
correspondingly---we can swap the module in the list for overriding,
retain the module in the list for inheritance, and add new module into
the list for extension. The compilation from a family (internal
representation) to a module still acts the same.

Though inheritance is solely a first-order data rather
than a ``mapping'', it is still possible to
\text{mix-in} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance. However, a good definition of mixin is still under investigation. We will show one practical example of \textit{mixin} later.\YZ{How about making mixin composition one of your Req's? Also does any example in the paper illustrate how to do this?}\EDJreply{mixin is not really formalized. There is no example illustrate it. The only example is the STLC example at the very end and I just kind of ``mention'' it. I think here I just want to emphasize, inheritance as first order data is still possible to support composition.}

@@

% For example, we may
% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
% and intend to prove its commutativity in a following field $"comm"$.
% But to prove so, we need to know more than just a type interface of
% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
% future, once "add" is overridden into a subtraction function then
% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
% A more concrete example is in \cref{fig:plugin-example1}, consider the
% case we have "Field b : a = 1 := .." instead. Then our
% compilation and Coq's type checking will both fail: in the context of
% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
% more (and this is necessary so that "a" can be overridden in the
% future).
Take the field "c" of "Family I" in \cref{fig:plugin-example3} as the
example. We can see that the first attempt of proving "a = 1" fails.
\textit{This should fail}---because we want to retain the possibility
to override "a" in the future, but the mere existence of the field "c"
will stop that, since "c" witnesses "a" fix to "1". \textit{This will fail} in our implementation---the proof of "c" will require a concrete definition of "a",
but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in (a module with) the context where we only know \mintinline{Coq}{a : nat} thus failed.

@@

\textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.} 
Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming. There are cases where overriding and
(definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
because to prove
the latter we need to have concrete definition of "f" in the family type, which
makes "f" non-overridable at all. One option is to use sigma type
\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
a bundle of stuff with multiple refinements.

To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type. In a sealed family, every field will expose its definition inside the family but a family as a whole will be sealed into a ``relaxed'' type. Thus we can override one sealed family as a whole and change the internal implementation.

@@

And the compiled module type of
"MonotonicF" will be fixed throughout the extension in different
families. Any overriding family will always have these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }

In our example, the seal family "I.MonotonicF" is actually simulating
interfaces as well.  We have
\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
and we assigned it to both "f" and "p", and later instantiate in
"I2" with concrete implementation by overriding this family
"I.MonotonicF".
% Basically, {a : int -> int; property : a is monotonic}, in this case
%   we may want to override "a" using different computation, but we still wnat
%   this refinement/property
The implementation of sealed family is not so different from that of the normal family, but we need to be careful about the compiled module type.

@@

We start with the programming interface. We use "FInductive" to define an extensible inductive type and in the children family we use "Extend Finductive ..." to extend the type with new constructors. In \cref{fig:plugin-example2}, we make the example of boolean and three-valued boolean.

Recall, to construct a recursor, we need to create for each
constructor a handler (function) and aggregate them into a family (e.g. "neg_handler" in \cref{fig:plugin-example2}). We use "FRecursor" to declare a recursor (e.g. "neg") by specifying the motive, the aiming inductive type and the handler family. Once recursor is created it will look like any other function fields (e.g. "neg"). Here in \cref{fig:plugin-example2} we apply "neg" to "tt" as an exemplar field. The inheritance of recursor is mainly delegated by the inheritance of the handler family---if the user extend with new constructor (e.g. "uu"), then the inherited family (e.g. "neg_handler") will need to extend correspondingly (e.g. "neg_handler.uu"). If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".\EDJ{This paragraph is duplicating but I am not sure how to make it concise.}

@@

Let's look at the exemplar \cref{fig:plugin-example2}, the example of boolean and negation. When we define the extensible inductive type "b", all the following fields (e.g. "neg_handler.ff") can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface''---the compiled module type "b_3". We cannot export\YZ{export or put?}\EDJreply{Ok. I just figure out how my wording is confusing. put is absolutely correct but by put I want to mean ``export''. I think I use ``export'' in several places...}
the eliminator of the inductive type "b" to "b_3" because doing so all of the following fields (e.g. "neg_handler.ff" and the corresponding "ff_24") will not be able to be inherited to the context where "b"
is extended with a third constructor, \textbf{since the mere existence of the recursor will assert that there are only two constructors} in "b", not
\textbf{extensible} at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}\EDJreply{You are right. It is not only about "ff_24".I rephrased it. Please check.}

However, we later still need the concrete eliminator "b" for doing exhaustiveness checking (as in \ref{chg:extensible-inductive-type}), thus we will store the corresponding info somewhere.  Here we show one exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module "b_rec_12". This concrete eliminator will be used by our "FRecursor". For example, in \cref{fig:plugin-example2}, "FRecursor" command will feed Coq a module "v_33_34" for doing type-checking with specified handlers (e.g. "neg_handler") during compilation---what it does is basically
assert the existence of the recursor and apply to see if handlers can
pass the type checking by Coq. Note that, at this point of the example, "FRecursor" will only do type check. The compilation of "FRecursor" into a component of the compiled module only happens when closing the whole family, by inserting the following into the compiled module: 
\begin{minted}{Coq}
  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff.
\end{minted}
where "b_rec" and "b" are the corresponding recursor and the (vanilla) inductive type in that compiled module. Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields.

Importantly, the recursor field (e.g. "neg") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) modules, a recursor field is not compiled like that. For all the children family, we have to reconstruct and override the "FRecursor" field with a ``new'' recursor using (maybe the same) handler family. This is affordable since each aggregated recursor handler doesn't need
to be rechecked---look at "B2.neg_handler", there is only one new
handler for "uu" and the handlers (mundane fields) from the parent
"B.neg_handler.tt,B.neg_handler.ff" have already been type-checked and
inherited (since they only rely on ``the abstracted
interface'' module "b_3"). 
Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg". For example, for the new "neg", when compiling family "B2" into a module, at the appropriate place our plugin will insert
\begin{minted}{Coq}
  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff neg_handler.uu.
\end{minted}
where "b_rec" and "b" are the corresponding recursor and the (vanilla but extended and modified) inductive type in that compiled module. Note that, even though "b_rec" and "b" has the same name as before, they are in the newly-compiled module so they refer to totally different things.

Of course, in the implementation, our plugin will ``inherit'' the
recursor "neg" in the surface \YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}\EDJreply{Added above. Please Check.}
but what the plugin actually does is doing a second exhaustiveness
checking in children family like mentioned above, with the ``same'' motive and the ``same'' handler family.\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea. Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}\EDJreply{Please check.}

@@

% explain ftheorem wrapping this complicated recursor
This design of decoupling of the case handlers and exhaustiveness checking handles the semantic well, however, it brings a lot of boilerplate code---we need to create an extra (handler) family, manually specifying the type of each case handlers, and then using "FRecursor" to ``tie the knot'' and construct a recursor field from the handlers.  
We thus provide a command "FTheorem" to avoid these boilerplate in programming interface when using tactic programming. 
This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor".\YZ{example?}\EDJreply{FTheorem is really just a short hand for "FRecursor" + handler family. Nothing more conceptual here. If the reader is interested in the concrete syntax, I think I need to make the two examples avaliable in the appendix.} We expect this "FTheorem" to be used in theorem proving but not general programming.

To summarize, (1) to make each recursor
handler inheritable, we need to seal an abstraction around the inductive
type (e.g., module "b_3"); but to construct a recursor (and carry out the
exhaustiveness checking), we need to break this abstraction and see the
concrete definition of the inductive type (e.g., module "b_rec_12"). Our
meta-theory need to handle these two seemingly contradicting ideas
simultaneously when solving \ref{chg:extensible-inductive-type} . (2) \ref{chg:extensible-inductive-type} is also resolved by this \textit{decoupling} of the syntax of implementation of recursor handlers and exhaustiveness checking. The former is handled by family inheritance and thus avoiding the boilerplate code; and the non-inheritance of the compiled recursor ensures the exhaustiveness checking happens for every recursor.
